SafeNet: A European database of plant-pollinator interactions

### 2024-03-06 Manuscript/Figure1.qmd
Table with data sturcture added to MS

#Load data
data = readRDS("../Data/Interactions_uncounted.rds")

@@

#Plot
p1 = ggplot(poll_spread, 
aes(reorder(Pollinator_accepted_name,-Percent_total), Percent_total)) +
geom_bar(stat = "identity", 
           color = "#D55E00",
           lwd = 0.015, fill = "#D55E00") +
ylab("Shared pollinators (%)") +
theme_minimal()+
theme(axis.text.x = element_blank(),
  axis.ticks.x=element_blank(),
  axis.ticks.y = element_line(size = 0.2),
  plot.title = element_text(size = 8, face = "bold"), 
  axis.text.y = element_text(size=6),
  panel.grid.major.x = element_blank(),
  panel.grid.minor.y = element_blank(),
  panel.grid.major.y = element_line(size=0.2, linetype =           2)) +
xlab(NULL) +
ylab(NULL) + 
ggtitle("Pollinators")+
geom_vline(xintercept = quantile_50, linetype="dashed", 
color = "black", size=0.25) + 
annotate(geom="text", x=1285, y=0.83, label="50th percentile",          color="black", size = 1.5) +
annotate(geom="text", x=20, y=0.88, label=most_common_poll[1],         color="black",  fontface = 'italic', size = 1.5)+
scale_y_continuous(breaks = seq(0, 0.75, by= 0.25), expand = c(0, 0)) + 
annotate(geom="text", x=-500, y=1.15, label="(b)",color="black", size = 3.5) +
coord_cartesian(xlim=c(0, nlevels(factor(poll_spread$Pollinator_accepted_name))),ylim=c(0, max(poll_spread$Percent_total)), clip="off")

@@

#Plot
p2 = ggplot(plant_spread, 
aes(reorder(Plant_accepted_name,-Percent_total), Percent_total, fill = Plant_accepted_name)) +
geom_bar(stat = "identity", 
           color = "#009E73",
           lwd = 0.015, fill = "#009E73") +
ylab("Shared plants (%)") +
theme_minimal()+
theme(axis.text.x = element_blank(),
         axis.ticks.x=element_blank(),
         axis.ticks.y = element_line(size = 0.2),
         axis.title.x = element_text(face = "bold", size = 8),            plot.title = element_text(size = 8, face = "bold"),
         axis.text.y = element_text(size=6),
         panel.grid.major.x = element_blank(),
         panel.grid.minor.y = element_blank(),
         panel.grid.major.y = element_line(size=0.2, linetype =           2)) +
xlab("Species") +
ylab(NULL) + 
ggtitle("Plants") + 
geom_vline(xintercept = quantile_50, linetype="dashed", 
          color = "black", size=0.25) + 
annotate(geom="text", x=835, y=0.7, label="50th percentile",color="black", size = 1.5) + 
annotate(geom="text", x=20, y=0.75, label=most_common_plant[1]
      ,color="black",  fontface = 'italic', size = 1.5)+
annotate(geom="text", x=20, y=-0.08, label="Common spp."
      ,color="black",  size = 1.5) +
annotate(geom="text", x=1200, y=-0.08, label="\'Rare\' spp."
      ,color="black",  size = 1.5) + 
scale_y_continuous(breaks = seq(0, 0.6, by= 0.2), expand = c(0, 0), labels = label_number(accuracy = 0.01)) +
coord_cartesian(xlim=c(0, nlevels(factor(plant_spread$Plant_accepted_name))),ylim=c(0, max(plant_spread$Percent_total)), clip="off")

#Prepare panel with plots
plot2 =  wrap_elements(p1 / p2) +  labs(tag ="Occurrence across studies (%)") + 
theme(plot.tag = element_text(size = 7, angle = 90, face = "bold", vjust = -3.5), plot.tag.position = "left")
#Plot everything
map + plot_spacer() + plot2 + plot_layout(widths = c(2.1, -0.45, 1))

>>>>

#Load data
data = readRDS("../Data/3_Final_data/Interactions_uncounted.rds")

@@

#Plot
p1 = ggplot(poll_spread, 
aes(reorder(Pollinator_accepted_name,-Percent_total), Percent_total)) +
geom_bar(stat = "identity", 
           color = "#D55E00",
           lwd = 0.015, fill = "#D55E00") +
ylab("Shared pollinators (%)") +
theme_minimal()+
theme(axis.text.x = element_blank(),
  axis.ticks.x=element_blank(),
  axis.ticks.y = element_line(size = 0.2),
  plot.title = element_text(size = 8, face = "bold"), 
  axis.text.y = element_text(size=5),
  panel.grid.major.x = element_blank(),
  panel.grid.minor.y = element_blank(),
  panel.grid.major.y = element_line(size=0.2, linetype =           2)) +
xlab(NULL) +
ylab(NULL) + 
ggtitle("Pollinators")+
geom_vline(xintercept = quantile_50, linetype="dashed", 
color = "black", size=0.25) + 
annotate(geom="text", x=1285, y=0.83, label="50th percentile",          color="black", size = 1.5) +
annotate(geom="text", x=20, y=0.88, label=most_common_poll[1],         color="black",  fontface = 'italic', size = 1.5)+
scale_y_continuous(breaks = seq(0, 0.75, by= 0.25), expand = c(0, 0)) + 
annotate(geom="text", x=-500, y=1.15, label="(c)",color="black", size = 3.5) +
coord_cartesian(xlim=c(0, nlevels(factor(poll_spread$Pollinator_accepted_name))),ylim=c(0, max(poll_spread$Percent_total)), clip="off")

@@

#Plot
p2 = ggplot(plant_spread, 
aes(reorder(Plant_accepted_name,-Percent_total), Percent_total, fill = Plant_accepted_name)) +
geom_bar(stat = "identity", 
           color = "#009E73",
           lwd = 0.015, fill = "#009E73") +
theme_minimal()+
theme(axis.text.x = element_blank(),
         axis.ticks.x=element_blank(),
         axis.ticks.y = element_line(size = 0.2),
         axis.title.x = element_text(face = "bold", size = 7),            plot.title = element_text(size = 8, face = "bold"),
         axis.text.y = element_text(size=5),
         panel.grid.major.x = element_blank(),
         panel.grid.minor.y = element_blank(),
         panel.grid.major.y = element_line(size=0.2, linetype =  2),
      axis.title.y = element_text(face= "bold", size= 7, hjust= -0.01, vjust = 1)) +
xlab("Species") +
ylab("Occurrence across studies") + 
ggtitle("Plants") + 
geom_vline(xintercept = quantile_50, linetype="dashed", 
          color = "black", size=0.25) + 
annotate(geom="text", x=835, y=0.7, label="50th percentile",color="black", size = 1.5) + 
annotate(geom="text", x=20, y=0.75, label=most_common_plant[1]
      ,color="black",  fontface = 'italic', size = 1.5)+
annotate(geom="text", x=20, y=-0.08, label="Common spp."
      ,color="black",  size = 1.5) +
annotate(geom="text", x=1200, y=-0.08, label="\'Rare\' spp."
      ,color="black",  size = 1.5) + 
scale_y_continuous(breaks = seq(0, 0.6, by= 0.2), expand = c(0, 0), labels = label_number(accuracy = 0.01)) +
coord_cartesian(xlim=c(0, nlevels(factor(plant_spread$Plant_accepted_name))),ylim=c(0, max(plant_spread$Percent_total)), clip="off")

data2 = data %>%  select(Study_id, Year) %>% 
distinct() %>% 
group_by(Year) %>% 
summarise(Counts = n())

p3 = ggplot(data2, aes(Year, Counts)) +
geom_col(fill = "gray25", width = 0.9, color="black", linewidth = 0.4) +
theme_light() +
theme(plot.margin=unit(c(0,0,0,0), "mm"),
  axis.title.y = element_text(face = "bold", size = 7, vjust = -1.1),
  axis.title.x = element_text(face = "bold", size = 8),
  axis.text.y = element_text(size = 6),
  axis.text.x = element_text(size = 6,  angle = 90), 
  strip.text.x = element_text(size=12),
   axis.ticks.x = element_line(color="black"),
  strip.background = element_rect(
  color="black", fill=NA, size=0.5, linetype="solid")) + 
ylab("Number of studies") +
coord_cartesian(expand = F, clip = "off") + 
theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
  panel.background = element_blank(),
  panel.border = element_rect(colour = "black", fill=NA, size=1)) +
theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
panel.background = element_blank(),
panel.border = element_blank(),
axis.line = element_line(color = 'black')) +
scale_y_continuous(breaks = c(0, 4,8,12), labels = c("0", "4","8", "12"), limits = c(0,12)) +
scale_x_continuous(breaks = seq(min(data2$Year), max(data2$Year), by = 2), limits = c(min(data2$Year),max(data2$Year)))+
ggtitle("(b)") +
theme(plot.title = element_text(size = 10,hjust = -0.35, vjust = 0.8))

plot2 = p3 / p1 / p2

#Plot everything
map + plot_spacer() + plot2 + plot_layout(widths = c(2.1, -0.45, 0.8))

### 2024-03-06 Manuscript/Figure_4.Rmd
Table with data sturcture added to MS

```{r echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE, out.width="95%", fig.cap= "\\textbf{Figure 2}. Proportion of the major pollinator orders by habitat types and bioclimatic regions in the SafeNet database. The orders, from left to right, include Hymenoptera, Diptera, Lepidoptera and Coleoptera. The horizontal barplot on the right indicates the number of studies that were conducted on each habitat type or bioclimatic region. Note that a single study can countribute to more than one habitat or bioclimatic region. Areas with a greater number of studies are more likely to depict accurate proportions of pollinators in those systems.", fig.width=6, fig.height=8}

@@

p1 = ggplot(c2, 
aes(fill=Pollinator_order, y=Proportion, x=Habitat)) + 
geom_bar(position = position_stack(reverse = TRUE), stat="identity")+
theme_bw() +
theme(legend.position = "bottom",
      plot.margin=unit(c(0,0,0,0), "mm"),
      legend.text = element_text(size=6),
      legend.key.size = unit(2, "mm"),
      legend.spacing.y = unit(0.1, "cm"),
       legend.box.spacing = unit(0, "pt"),
      axis.title = element_text(face = "bold", size = 8),
      axis.text = element_text(size=6),
      panel.border = element_rect(size=1)) +
xlab(NULL) +
scale_y_continuous(breaks = c(0.25, 0.5,0.75), labels = c("0.25", "0.5","0.75"))+
theme(strip.text.x = element_text(size = 12, face = "bold"),
    strip.background = element_rect(
color=NA, fill=NA, size=0.5, linetype="solid")) +
scale_fill_manual(values = viridis(4),
                  name = NULL) +
coord_flip(expand = FALSE) +
ggforce::facet_col(~ factor(Group, levels=c("Habitat","Bioregion")), space = "free", scales = "free_y") +
guides(fill=guide_legend(nrow=2,byrow=TRUE)) +
ggtitle("(a)") +
theme(plot.title = element_text(size = 10, face = "bold",hjust = -0.25, vjust = 1))

@@

data2 = data %>%  select(Study_id, Year) %>% 
distinct() %>% 
group_by(Year) %>% 
summarise(Counts = n())

p3 = ggplot(data2, aes(Year, Counts)) +
geom_col(fill = "gray25", width = 0.9, color="black", linewidth = 0.4) +
theme_light() +
theme(plot.margin=unit(c(0,0,0,0), "mm"),
  axis.title.y = element_text(face = "bold", size = 8, vjust = -2),
  axis.title.x = element_text(face = "bold", size = 8),
  axis.text.y = element_text(size = 6),
  axis.text.x = element_text(size = 6,  angle = 90), 
  strip.text.x = element_text(size=12),
   axis.ticks.x = element_line(color="black"),
  strip.background = element_rect(
  color="black", fill=NA, size=0.5, linetype="solid")) + 
ylab("Number of studies") +
coord_cartesian(expand = F, clip = "off") + 
theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
  panel.background = element_blank(),
  panel.border = element_rect(colour = "black", fill=NA, size=1)) +
theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
panel.background = element_blank(),
panel.border = element_blank(),
axis.line = element_line(color = 'black')) +
scale_y_continuous(breaks = c(0, 4,8,12), labels = c("0", "4","8", "12"), limits = c(0,12)) +
scale_x_continuous(breaks = seq(min(data2$Year), max(data2$Year), by = 2), limits = c(min(data2$Year),max(data2$Year)))+
ggtitle("(b)") +
theme(plot.title = element_text(size = 10, face = "bold",hjust = -0.25, vjust = 1))

a = plot_spacer() +  p3 + plot_spacer() + plot_layout(ncol=1, heights= c(9,10,1))
#Plot panel
p1 + p2 + a +plot_layout(widths = c(4,2, 2.5),heights= c(10,10,4), ncol=3)

>>>>

```{r echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE, out.width="95%", fig.cap= "\\textbf{Figure 2}. Proportion of the major pollinator orders by habitat types and bioclimatic regions in the SafeNet database. The orders, from left to right, include Hymenoptera, Diptera, Lepidoptera and Coleoptera. The horizontal barplot on the right indicates the number of studies that were conducted on each habitat type or bioclimatic region. Note that a single study can countribute to more than one habitat or bioclimatic region. Areas with a greater number of studies are more likely to depict accurate proportions of pollinators in those systems.", fig.width=4.5, fig.height=8, fig.align='center'}

@@

p1 = ggplot(c2, 
aes(fill=Pollinator_order, y=Proportion, x=Habitat)) + 
geom_bar(position = position_stack(reverse = TRUE), stat="identity")+
theme_bw() +
theme(legend.position = "bottom",
      plot.margin=unit(c(0,0,0,0), "mm"),
      legend.text = element_text(size=6),
      legend.key.size = unit(2, "mm"),
      legend.spacing.y = unit(0.1, "cm"),
       legend.box.spacing = unit(0, "pt"),
      axis.title = element_text(face = "bold", size = 8),
      axis.text = element_text(size=6),
      panel.border = element_rect(size=1)) +
xlab(NULL) +
scale_y_continuous(breaks = c(0.25, 0.5,0.75), labels = c("0.25", "0.5","0.75"))+
theme(strip.text.x = element_text(size = 12, face = "bold"),
    strip.background = element_rect(
color=NA, fill=NA, size=0.5, linetype="solid")) +
scale_fill_manual(values = viridis(4),
                  name = NULL) +
coord_flip(expand = FALSE) +
ggforce::facet_col(~ factor(Group, levels=c("Habitat","Bioregion")), space = "free", scales = "free_y") +
guides(fill=guide_legend(nrow=2,byrow=TRUE))

@@

#Plot panel
p1 + p2 + plot_layout(widths = c(4,2, 2.5),heights= c(10,10,4), ncol=2)
